---
title: "CCAMLRGIS R Package"
author: 'CCAMLR Secretariat'
output: rmarkdown::html_vignette
date: "`r format(Sys.time(), '%d %b %Y')`"
vignette: >
  %\VignetteIndexEntry{CCAMLRGIS R Package}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message = FALSE}
library(CCAMLRGIS)
set.seed(1234)
```

### A package to load and create spatial data, including layers and tools that are relevant to CCAMLR activities.

# Introduction

The CCAMLRGIS package was developed to simplify the production of maps in the CCAMLR Convention Area. It provides two categories of functions: load functions and create functions. Load functions are used to import spatial layers from the online CCAMLR GIS (https://gis.ccamlr.org/) such as the ASD boundaries. Create functions are used to create layers from user data such as polygons and grids. All spatial manipulations are made using the South Pole Lambert Azimuthal Equal Area projection (type ?CCAMLRp for more details).

First, install the package by typing:
```{r eval=F}
install.packages("CCAMLRGIS")
```

Then, load the package by typing:
```{r eval=F}
library(CCAMLRGIS)
```

## 1. Basemaps

Prior to detailing the package's capabilities, a set of basic commands are shown here to display a few core mapping elements.

#### Circumpolar map:

```{r fig.align="center",out.width="100%",message=F,dpi=200,eval=F}
#Load ASDs and EEZs
ASDs=load_ASDs()
EEZs=load_EEZs()
#Set the figure margins as c(bottom, left, top, right)
par(mai=c(0,0.4,0,0))
#Plot the bathymetry
plot(SmallBathy,breaks=Depth_cuts,col=Depth_cols,legend=F,axes=F,box=F)
#Add color scale
CBar(height=65,fontsize=0.4,offset=600,width=25,maxVal=-1,lwd=0.5)
#Add reference grid
add_RefGrid(bb=bbox(SmallBathy),ResLat=10,ResLon=20,LabLon=0,fontsize=0.4,lwd=0.5)
#Add ASD and EEZ boundaries
plot(ASDs,add=T,lwd=0.5,border='red')
plot(EEZs,add=T,lwd=0.5,border='red')
#Add ASD labels
add_labels('ASDs',fontsize=0.3,col='red')
```


#### Local map (e.g. Subarea 48.6):

```{r fig.align="center",out.width="100%",message=F,dpi=200,eval=F}
#Load ASDs
ASDs=load_ASDs()
#Subsample ASDs to only keep Subarea 48.6
S486=ASDs[ASDs$GAR_Short_Label=='486',]
#Crop bathymetry to match the extent of S486
B486=raster::crop(SmallBathy,S486)
#Optional: mask the bathymetry outside of S486
#Try running this code with and without 'mask' to see the difference
B486=raster::mask(B486,S486)
#Optional: get the maximum depth in that area to constrain the color scale
minD=raster::minValue(B486)
#Set the figure margins as c(bottom, left, top, right)
par(mai=c(0,0.2,0,0))
#Plot the bathymetry
plot(B486,breaks=Depth_cuts,col=Depth_cols,legend=F,axes=F,box=F)
#Add color scale
CBar(height=65,fontsize=0.4,offset=600,width=23,lwd=0.5,minVal=minD,maxVal=-1)
#Add reference grid
add_RefGrid(bb=bbox(B486),ResLat=5,ResLon=10,fontsize=0.4,lwd=0.5,offset = 100000)
#Add Subarea 48.6 boundaries
plot(S486,add=T,lwd=0.5,border='red')
#Add a -2000m contour
raster::contour(B486,levels=-2000,add=T,lwd=0.5,labcex=0.3)
#Add single label at the centre of the polygon
l=rgeos::polygonsLabel(S486,labels='48.6',col='red')
```

## 2. Create functions

### 2.1. Points, lines, polygons and grids

These functions are used to transform user data into spatial layers with the appropriate projection. User data may either be generated within an R script or imported from a '.csv' file located in the working directory. Users can set their working directory using the command setwd(). It is however recommended to, instead, create an R project in a folder and put your '.csv' files in it.

To see your current working directory, type:

```{r eval=F}
getwd()
```

#### Create points:
For details, type:
```{r eval=F}
?create_Points
```

```{r fig.align="center",out.width="100%",message=F,dpi=200,eval=F}

#Prepare layout for 4 sub-plots
par(mfrow=c(2,2))
#Set the figure margins as c(bottom, left, top, right)
par(mai=c(0,0.01,0.2,0.01))

#Example 1: Simple points with labels
MyPoints=create_Points(PointData)
plot(MyPoints,main='Example 1',cex.main=0.5,cex=0.5,lwd=0.5)
text(MyPoints$x,MyPoints$y,MyPoints$name,adj=c(0.5,-0.5),xpd=T,cex=0.5)
box()

#Example 2: Simple points with labels, highlighting one group of points with the same name
MyPoints=create_Points(PointData)
plot(MyPoints,main='Example 2',cex.main=0.5,cex=0.5,lwd=0.5)
text(MyPoints$x,MyPoints$y,MyPoints$name,adj=c(0.5,-0.5),xpd=T,cex=0.5)
plot(MyPoints[MyPoints$name=='four',],bg='red',pch=21,cex=1,add=T)
box()

#Example 3: Buffered points with radius proportional to catch
MyPoints=create_Points(PointData,Buffer=1*PointData$Catch)
plot(MyPoints,col='green',main='Example 3',cex.main=0.5,cex=0.5,lwd=0.5)
text(MyPoints$x,MyPoints$y,MyPoints$name,adj=c(0.5,0.5),xpd=T,cex=0.5)
box()

#Example 4: Buffered points with radius proportional to catch and clipped to the Coast
MyPoints=create_Points(PointData,Buffer=2*PointData$Catch,Clip=T)
plot(MyPoints,col='cyan',main='Example 4',cex.main=0.5,cex=0.5,lwd=0.5)
plot(Coast,add=T,col='grey',lwd=0.5)
box()
```

#### Create lines:
For details, type:
```{r eval=F}
?create_Lines
```

```{r fig.align="center",out.width="100%",message=F,dpi=200,eval=F,fig.height=1}

#Prepare layout for 3 sub-plots
par(mfrow=c(1,3))
#Set the figure margins as c(bottom, left, top, right)
par(mai=c(0,0.01,0.2,0.01))

#Example 1: Simple and non-densified lines
MyLines=create_Lines(LineData)
plot(MyLines,col=rainbow(length(MyLines)),main='Example 1',cex.main=0.5,lwd=1)
box()

#Example 2: Simple and densified lines (note the curvature of the purple line)
MyLines=create_Lines(LineData,Densify=T)
plot(MyLines,col=rainbow(length(MyLines)),main='Example 2',cex.main=0.5,lwd=1)
box()

#Example 3: Densified, buffered and clipped lines
MyLines=create_Lines(LineData,Densify=T,Buffer=c(10,40,50,80,100),Clip=T)
plot(MyLines,col=rainbow(length(MyLines)),main='Example 3',cex.main=0.5,lwd=1)
plot(Coast,col='grey',add=T,lwd=0.5)
box()
```

#### Create polygons:
For details, type:
```{r eval=F}
?create_Polys
```

```{r fig.align="center",out.width="100%",message=F,dpi=200,eval=F,fig.height=1.5}

#Prepare layout for 3 sub-plots
par(mfrow=c(1,3))
#Set the figure margins as c(bottom, left, top, right)
par(mai=c(0,0.01,0.2,0.01))

#Example 1: Simple and non-densified polygons
MyPolys=create_Polys(PolyData,Densify=F)
plot(MyPolys,col='blue',main='Example 1',cex.main=0.5,lwd=0.5)
text(MyPolys$Labx,MyPolys$Laby,MyPolys$ID,col='white',cex=0.5)
box()

#Example 2: Simple and densified polygons (note the curvature of iso-latitude lines)
MyPolys=create_Polys(PolyData)
plot(MyPolys,col='red',main='Example 2',cex.main=0.5,lwd=0.5)
text(MyPolys$Labx,MyPolys$Laby,MyPolys$ID,col='white',cex=0.5)
box()

#Example 3: Buffered and clipped polygons
MyPolysBefore=create_Polys(PolyData,Buffer=c(10,-15,120))
MyPolysAfter=create_Polys(PolyData,Buffer=c(10,-15,120),Clip=T)
plot(MyPolysBefore,col='green',main='Example 3',cex.main=0.5,lwd=0.5)
plot(Coast,add=T,lwd=0.5)
plot(MyPolysAfter,col='orange',add=T,lwd=0.5)
text(MyPolysAfter$Labx,MyPolysAfter$Laby,MyPolysAfter$ID,col='white',cex=0.5)
box()
```

#### Create grids:
For details, type:
```{r eval=F}
?create_PolyGrids
```

```{r fig.align="center",out.width="100%",message=F,dpi=200,eval=F,fig.height=1}

#Prepare layout for 3 sub-plots
par(mfrow=c(1,3))
#Set the figure margins as c(bottom, left, top, right)
par(mai=c(0,0.01,0.2,0.01))

#Example 1: Simple grid, using automatic colors
MyGrid=create_PolyGrids(GridData,dlon=2,dlat=1)
plot(MyGrid,col=MyGrid$Col_Catch_sum,main='Example 1',cex.main=0.5,lwd=0.1)
box()

#Example 2: Equal area grid, using automatic colors
MyGrid=create_PolyGrids(GridData,Area=10000)
plot(MyGrid,col=MyGrid$Col_Catch_sum,main='Example 2',cex.main=0.5,lwd=0.1)
box()

#Example 3: Equal area grid, using custom cuts and colors
MyGrid=create_PolyGrids(GridData,Area=10000,cuts=c(0,50,100,500,2000,3500),cols=c('blue','red'))
plot(MyGrid,col=MyGrid$Col_Catch_sum,main='Example 3',cex.main=0.5,lwd=0.1)
box()

```

Customizing a grid and adding a color scale:

```{r fig.align="center",out.width="100%",message=F,dpi=200,eval=F,fig.height=3}

#Prepare layout for 2 sub-plots
par(mfrow=c(2,1))

#Step 1: Generate your grid
MyGrid=create_PolyGrids(GridData,Area=10000)

#Step 2: Inspect your gridded data (e.g. sum of Catch) to determine whether irregular cuts are required
par(mai=c(0.4,0.4,0.1,0.1)) #set plot margins as c(bottom, left, top, right)
hist(MyGrid$Catch_sum,100,cex=0.5,main='Frequency distribution of data',
     cex.main=0.5,lwd=0.2,col='grey',cex.axis=0.5)
box()
#In this case (heterogeneously distributed data) irregular cuts would be preferable
#Such as:
MyCuts=c(0,50,100,500,2000,2500)
abline(v=MyCuts,col='green',lwd=0.1,lty=2) #Add classes to histogram as green dashed lines

#Step 3: Generate colors according to the desired classes (MyCuts)
Gridcol=add_col(MyGrid$Catch_sum,cuts=MyCuts,cols=c('yellow','purple'))

#Step 4: Plot result and add color scale
par(mai=c(0,0,0,0.8)) #set plot margins as c(bottom, left, top, right)
#Use the colors generated by add_col
plot(MyGrid,col=Gridcol$varcol,lwd=0.1) 
#Add color scale using cuts and cols generated by add_col
CBar(title='Sum of Catch (t)',Bcuts=Gridcol$cuts,Bcols=Gridcol$cols,width=26,
     fontsize=0.35,lwd=0.5) 
box()

```

### 2.2. Create Stations

This function was designed to create random point locations inside a polygon and within bathymetry strata constraints. A distance constraint between stations may also be used if desired.

For details, type:
```{r eval=F}
?create_Stations
```

First, create a polygon within which stations will be created:

```{r fig.align="center",out.width="100%",message=F,dpi=200,eval=T}
#Create polygons
MyPolys=create_Polys(PolyData,Densify=T)

#Set the figure margins as c(bottom, left, top, right)
par(mai=c(0,0,0,0))
plot(MyPolys)

#Subsample MyPolys to only keep the polygon with ID 'one'
MyPoly=MyPolys[MyPolys$ID=='one',]

plot(MyPoly,col='green',add=T)
text(MyPolys$Labx,MyPolys$Laby,MyPolys$ID)

```

Example 1. Set numbers of stations, no distance constraint:

```{r fig.align="center",out.width="100%",message=F,dpi=200,eval=F}

#Create polygon as shown above
MyPolys=create_Polys(PolyData,Densify=T)
MyPoly=MyPolys[MyPolys$ID=='one',]

#optional: crop your bathymetry raster to match the extent of your polygon
BathyCroped=raster::crop(SmallBathy,MyPoly)

MyStations=create_Stations(MyPoly,BathyCroped,Depths=c(-550,-1000,-1500,-2000),N=c(20,15,10))
#Set the figure margins as c(bottom, left, top, right)
par(mai=c(0,0,0,0)) 

plot(BathyCroped,breaks=Depth_cuts,col=Depth_cols,legend=F,axes=F,box=F)
CBar(height=90,fontsize=0.5,width=25,lwd=0.5,offset=0,maxVal=-1)
plot(MyPoly,add=T,border='red',lwd=1)
contour(BathyCroped,levels=c(-550,-1000,-1500,-2000),add=T)
plot(MyStations,add=T,col='orange',cex=0.5)
```


Example 2. Set numbers of stations, with distance constraint:

```{r fig.align="center",out.width="100%",message=F,dpi=200,eval=F}

#Create polygon as shown above
MyPolys=create_Polys(PolyData,Densify=T)
MyPoly=MyPolys[MyPolys$ID=='one',]

#optional: crop your bathymetry raster to match the extent of your polygon
BathyCroped=raster::crop(SmallBathy,MyPoly)

MyStations=create_Stations(MyPoly,BathyCroped,Depths=c(-550,-1000,-1500,-2000),N=c(20,15,10),dist=10)
#Set the figure margins as c(bottom, left, top, right)
par(mai=c(0,0,0,0)) 

plot(BathyCroped,breaks=Depth_cuts,col=Depth_cols,legend=F,axes=F,box=F)
CBar(height=90,fontsize=0.5,width=25,lwd=0.5,offset=0,maxVal=-1)
plot(MyPoly,add=T,border='red',lwd=1)
contour(BathyCroped,levels=c(-550,-1000,-1500,-2000),add=T)
plot(MyStations[MyStations$Stratum=='550-1000',],pch=21,bg='yellow',add=T,cex=0.75,lwd=0.1)
plot(MyStations[MyStations$Stratum=='1000-1500',],pch=21,bg='orange',add=T,cex=0.75,lwd=0.1)
plot(MyStations[MyStations$Stratum=='1500-2000',],pch=21,bg='red',add=T,cex=0.75,lwd=0.1)
```


Example 3. Automatic numbers of stations, with distance constraint:

```{r fig.align="center",out.width="100%",message=F,dpi=200,eval=F}

#Create polygon as shown above
MyPolys=create_Polys(PolyData,Densify=T)
MyPoly=MyPolys[MyPolys$ID=='one',]

#optional: crop your bathymetry raster to match the extent of your polygon
BathyCroped=raster::crop(SmallBathy,MyPoly)

MyStations=create_Stations(MyPoly,BathyCroped,Depths=c(-550,-1000,-1500,-2000),Nauto=30,dist=10)
#Set the figure margins as c(bottom, left, top, right)
par(mai=c(0,0,0,0)) 

plot(BathyCroped,breaks=Depth_cuts,col=Depth_cols,legend=F,axes=F,box=F)
CBar(height=90,fontsize=0.5,width=25,lwd=0.5,offset=0,maxVal=-1)
plot(MyPoly,add=T,border='red',lwd=1)
contour(BathyCroped,levels=c(-550,-1000,-1500,-2000),add=T)
plot(MyStations[MyStations$Stratum=='550-1000',],pch=21,bg='yellow',add=T,cex=0.75,lwd=0.1)
plot(MyStations[MyStations$Stratum=='1000-1500',],pch=21,bg='orange',add=T,cex=0.75,lwd=0.1)
plot(MyStations[MyStations$Stratum=='1500-2000',],pch=21,bg='red',add=T,cex=0.75,lwd=0.1)
```

## 3. Load functions

Download the up-to-date spatial layers from the online CCAMLRGIS and load them to your environment.

For details, type:
```{r eval=F}
?load_ASDs
```

```{r fig.align="center",out.width="100%",message=F,dpi=200,eval=F}

#Load ASDs, EEZs, and Coastline
ASDs=load_ASDs()
EEZs=load_EEZs()
Coastline=load_Coastline()

#Set the figure margins as c(bottom, left, top, right)
par(mai=c(0,0,0,0)) 
#Plot
plot(ASDs,col='green',border='blue')
plot(EEZs,col='orange',border='purple',add=T)
plot(Coastline,col='grey',add=T)
add_labels('ASDs',fontsize=0.5,col='red')

```

### Offline use

Since the 'load_' functions require an internet connection, users may desire to save layers on their hard drive for offline use. This may be done, at the risk of not having the most up-to-date layers, as follows:

```{r eval=F}

#Load all layers
ASDs=load_ASDs()
EEZs=load_EEZs()
Coastline=load_Coastline()
SSRUs=load_SSRUs()
RBs=load_RBs()
SSMUs=load_SSMUs()
MAs=load_MAs()
RefAreas=load_RefAreas()
MPAs=load_MPAs()

#Save as .RData file
save(list=c('ASDs','EEZs','Coastline','SSRUs','RBs','SSMUs','MAs','RefAreas','MPAs'),
     file = "CCAMLRLayers.RData")

#Later, when offline load layers:
load("CCAMLRLayers.RData")
```

## 4. Other functions

### 4.1. get_depths

### 4.2. seabed_area

### 4.3. assign_areas

## 5. Adding colors and legends


